---
title: "Getting Started"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To start using biophysical models, you must first install R on your computer. Go to `r-project.org` and install the software. However, the R-base has quite a few graphic devices and, therefore, an integrated development environment such as RStudio is strongly recommended. Go to `rstudio.com` and follow the installation instructions.

The next step is to install the `bproce` package. A package in R is a collection of functions that can be used by loading them into the R environment. In `bproce` package, each function implements a specific biophysical models. In order to get access to these functions, please make sure that the `bproce` package has been installed.

```{r eval=FALSE}
# install the 'bproce' package locally
system("R CMD INSTALL PATH_TO_PACKAGE/bproce_0.0.0.9000.tar.gz")
```

where `PATH_TO_PACKAGE` is the path of the package file `bproce_0.0.0.9000.zip` which is to be installed.

Load the package through the `require()` function:

```{r comment=""}
require(bproce)
```

To get help about R functions, use the `"?"` operator which provides access to documentation.

```{r eval=FALSE}
?hourlyT  #Find information about the 'hourlyT' function
```

You will find a documentation titled "Calculate Hourly Air Temperature". Inside the documentation there are several sections:

-   **Description:** information about what the function does.

-   **Usage:** the syntax of the function.

-   **Arguments:** a short description of the arguments used by the function.

-   **Value:** the object returned by the function.

-   **Examples:** examples on how to use the function.

If you are using RStudio, simply click the 'Run examples' link in the Examples section, and you will see the calculated results generated by the function `hourlyT()`. This action is exactly the same as the implementation of the following script.

```{r comment=""}
### ** Examples for calculating diel air temperature

lng <- 85.446758
lat <- 42.476708
t_max <- 30.2
t_min <- 14.3
doy <- DOY(2016, 9, 22)
hours <- seq(0, 23.5, 0.5)            # half-hour time resolution
time_sr <- sunRS(doy, lng, lat)[[1]]  # calculate sunrise time
time_ss <- sunRS(doy, lng, lat)[[2]]  # calculate sunset time
hourlyT(hours, t_max, t_min, time_sr, time_ss,
        alpha = 1.5, beta = 4, gamma = 1)
```

Section 1-1A will give more details about calculating diel air temperatures from theoretical perspectives.
