---
title: "Calculate Extraterrestrial Radiation"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate extraterrestrial radiation.

## Code Example

```{r eval=FALSE}
# Calculate Extraterrestrial Radiation
# ********************************
#
# Calculate daily extraterrestrial radiation (MJ m-2 d-1) (Allen et al., 1998).
#
# ********************************
# List of variables and units for this example
#
# 'year'        experimental year
# ’month'       experimental month
# 'day'         experimental day
# 'lat'         latitude of the study site (decimal)
# 
# Calculated variables in the program
#
# 'doy'         day of the year
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
year  <- 2016
month <- 9
day   <- 22
lat   <- 42.476708

# Step 3: run the model
doy <- DOY(year, month, day)
ra(lat, doy)
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `day`, `month`, and `year`: Experimental day. The provided example calculates the extraterrestrial radiation for 22 September 2016 at 42.476708^o^N and, therefore, we assigns 22 to the variable `day`, 9 to `month`, 2016 to `year`.

-   `lat`: Latitude of the study site (in decimal). The provided example focuses on extraterrestrial radiation at 42.476708^o^N.

```{r comment=""}
year  <- 2016
month <- 9
day   <- 22
lat   <- 42.476708
```

**Step3:** Run the model. The first step is to calculate day of the year (DOY) using the following equation:

$$
\text{DOY}=\left[ \frac{275\text{month}}{9} \right]-\left[\frac{\text{month}+9}{12}\right]·
\left(\left[ \frac{\text{year}-4\text[\text{year}/4]+2}{3} \right]+1\right)+\text{day}-30,\quad(1)
$$

```{r comment=""}
doy <- DOY(year, month, day)
cat(paste0(day,"/",month,"/",year," is the ",doy,"th day of ",year,".\n"))
```

The extraterrestrial radiation ($R_a$, MJ m^-2^ d^-1^) can be calculated as (Allen et al., 1998):

$$
R_a=\frac{24·60}{\pi}·G_{sc}d_r[\omega_s\sin\phi\sin\delta+\cos\phi\cos\delta\sin\omega_s],\quad(2)
$$ where $G_{sc}$ is solar constant (0.0820 MJ m^-2^ min^-1^), $d_r$ is inverse relative distance Earth-Sun, $\omega_s$ is sunset hour angle, $\phi$ is latitude (in radius), $\delta$ is solar decimation (in radius). Radius will be automatically converted into decimals through the function `ra`. $d_r$ is given by:

$$
d_r=1+0.033\cos\left( \frac{2\pi}{365}\text{DOY} \right),\quad(3)
$$

while $\delta$ can be calculated as:

$$
\delta=0.409\sin\left( \frac{2\pi}{365}\text{DOY}-1.39 \right), (4)
$$

The sunset hour angle $\omega_s$ is given by:

$$
\omega_s=\arccos\left( -\tan\phi\tan\delta\right),\quad(5)
$$

```{r comment=""}
Ra <- ra(lat, doy)

cat(paste0(
  "Date = ", day, "/", month, "/", year, "\n",
  "Latitude = ", lat, "\n",
  "Extraterrestrial Radiation = ", round(Ra, 3), " (MJ m-2 d-1)\n"
))
```

## Reference

Allen, R. G., Pereira, L. S., Raes, D. and Smith, M. (1998). Crop Evapotranspiration -- Guidelines for Computing Crop Water Requirements, FAO Irrigation and Drainge Paper 56, FAO, 1998, ISBN 92-5-104219-5.
