---
title: "Sunrise and Sunset Time"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate sunrise and sunset time through the function `sunRS()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate Sunrise and Sunset Time
# ********************************
#
# Calculate the sunrise and sunset time for a single day. For more details about 
# the algorithm used, see Campbell and Norman (2012).
#
# ********************************
# List of variables and units for this example
#
# 'year'        experimental year
# ’month'       experimental month
# 'day'         experimental day
# 'lng'         longitute of the study site (decimal)
# 'lat'         latitude of the study site (decimal)
# 
# Calculated variables in the program
# 'doy'         day of the year
# 'time_sr'     sunrise time (hour)
# 'time_ss'     sunset time (hour)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
year  <- 2016
month <- 9
day   <- 22
lng   <- -85.446758
lat   <- 42.476708

# Step 3: run the model
doy <- DOY(year, month, day)          # calculate day of the year (DOY)

time_sr <- sunRS(doy, lng, lat)[[1]]  # calculate sunrise time
time_ss <- sunRS(doy, lng, lat)[[2]]  # calculate sunset time

# Step 4: present the output
cat(
  paste0("Longitude = ",lng, ", Latitude = ",lat,".\n"),
  paste0("The sunrise time on ",day,"/",month,"/",year," is ",
         floor(time_sr),":",round((time_sr-floor(time_sr))*60),".\n"),
  paste0("The sunset time on ",day,"/",month,"/",year," is ",
         floor(time_ss),":",round((time_ss-floor(time_ss))*60),".\n"),
  "Day length =",round(time_ss-time_sr,2),"(hour).\n"
)
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `day`, `month`, and `year`: Experimental day. The provided example calculates sunrise and sunset time for 22 September 2016 at 42.476708^o^N and, therefore, we assigns 22 to the variable `day`, 9 to `month`, 2016 to `year`.

-   `lng`: Longitude of the study site (in decimal). The provided example focuses on the sunrise and sunset time at 85.446758^o^W. Thus we assign --85.446758 to `lng`.

-   `lat`: Latitude of the study site (in decimal). The provided example focuses on the sunrise and sunset time at 42.476708^o^N. Thus we assign 42.476708 to `lat`.

```{r comment=""}
year  <- 2016
month <- 9
day   <- 22
lng   <- -85.446758
lat   <- 42.476708
```

**Step3:** Run the model. The first step is to calculate day of the year (DOY) using the following equation:

$$
\text{DOY}=\left[ \frac{275\text{month}}{9} \right]-\left[\frac{\text{month}+9}{12}\right]·
\left(\left[ \frac{\text{year}-4\text[\text{year}/4]+2}{3} \right]+1\right)+\text{day}-30,\quad(1)
$$

```{r comment=""}
doy <- DOY(year, month, day)           # calculate day of the year (DOY)
cat(paste0(day,"/",month,"/",year," is the ",doy,"th day of ",year,".\n"))
```

Sunrise time (Time~sr~, hour) is calculated as:

$$
\text{Time}_{sr}=12-\frac{1}{15}\cos^{-1}\left(\frac{-\sin\Phi\sin\delta}{\cos\Phi\cos\delta} \right)-\frac{t}{60},\quad(2)
$$

where $\Phi$ is latitude (decimal), $\delta$ is solar declination (decimal), $t$ is local time (hour) which can be estimated from longitude. Sunset time (Time~ss~, hour) can be estimated by:

$$
\text{Time}_{ss}=12+\frac1{15}\cos^{-1}\left(\frac{-\sin\Phi\sin\delta}{\cos\Phi\cos\delta}\right)-\frac t{60},\quad(3)
$$

Solar declination ($\delta$) is calculated as:

$$
\delta = -23.25·\cos\left[\frac{360}{365}(\text{DOY}+10) \right],\quad(4)
$$

Calculation of Time~sr~ and Time~ss~ via equation (2--4) can be done through the function `sunRS`:

```{r comment=""}
time_sr <- sunRS(doy, lng, lat)[[1]]   # calculate sunrise time
time_ss <- sunRS(doy, lng, lat)[[2]]   # calculate sunset time

cat(
  paste0("Longitude = ",lng, ", Latitude = ",lat,".\n"),
  paste0("The sunrise time on ",day,"/",month,"/",year," is ",
         floor(time_sr),":",round((time_sr-floor(time_sr))*60),".\n"),
  paste0("The sunset time on ",day,"/",month,"/",year," is ",
         floor(time_ss),":",round((time_ss-floor(time_ss))*60),".\n"),
  "Day length =",round(time_ss-time_sr,2),"(hour).\n"
)
```

## Reference

Campbell, G. S., and Norman, J. (2012). An Introduction to Environmental Biophysics. Springer Science & Business Media. 286pp.
