---
title: "Sunrise and Sunset Time of a Period"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate sunrise and sunset time of a period through the function `sunRS()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate 2016 Sunrise and Sunset Time of Kellogg Biological Station
# ********************************
#
# Calculate the sunrise and sunset time for year 2016 at Kellogg Biological Station
# (KBS), Michigan, USA. For more details about the algorithm used, see Campbell 
# and Norman (2012).
#
# ********************************
# List of variables and units for this example
#
# 'path'        path of a CSV file, which conains one column (with table titles)
#                * column 1: observation date. Required format "dd/mm/yyyy"
# 'hours'       hours of the day
# 'lng'         longitute of the study site (decimal)
# 'lat'         latitude of the study site (decimal)
# 
# Calculated variables in the program
# 'doy'         day of the year
# 'time_sr'     sunrise time (hour)
# 'time_ss'     sunset time (hour)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
path  <- "./data/1-1B Switchgrass Time Series.csv"
lng   <- -85.446758
lat   <- 42.476708

# Step 3: read data
dat   <- read.csv(path)              # read the CSV file
                                     #  * column 1 ($date): observation date
                                     #  * column 2 ($t_max): not used
                                     #  * column 3 ($t_min): not used
day   <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[1]))
                                     # extract day numbers from characters
month <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[2]))
                                     # extract month numbers from characters
year  <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[3]))
                                     # extract year numbers from characters

# Step 4: run the model
doys <- DOY(year, month, day)        # calculate day of the year (DOY)

time_sr <- as.numeric(lapply(lapply(doys, sunRS, lng, lat), function(xx) xx[1]))
                                     # calculate sunrise time
time_ss <- as.numeric(lapply(lapply(doys, sunRS, lng, lat), function(xx) xx[2]))
                                     # calculate sunset time
res <- data.frame(
  date = dat$date, time_sr, time_ss
)                             # calculated result
                              #  * column 1  ($date): observation date
                              #  * column 2  ($time_sr): sunrise time
                              #  * column 3  ($time_ss): sunset time

# Step 5: show and visualize results
print(res, digits = 3)        # present the output

plot(
  doys, time_sr, col=2, ylim=range(c(time_sr, time_ss)), type="l",
  xlab = "Day of the year", ylab = "Temperature",
  main = "2016 Sunrise and Sunset Time of Kellogg Biological Station"
)                             # plot dew point temperature versus time
lines(doys, time_ss, col=3, type="l")
                              # plot wet-bulb temperature versus time
legend("right", pch=c(15, 15), col=c(2, 3), bty="n",
       legend=c("Sunrise Time", "Sunset Time")) # add legends
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `path`: the path of a CSV file which the data are to be read from. Each row of the file appears as one line, and columns are separated by commas (,). The first row appears as the table title. There should be three columns in this file:

    -   Column 1: Observation date. The format should be 'dd/mm/yyyy', where 'dd' refers to day, 'mm' refers to month and 'yyyy' refers to the year number. Even though the sample file "1-1B Switchgrass Time Series.csv" contains three columns, the last two columns are not used in this example.

-   `lng`: Longitude of the study site (in decimal). The provided example focuses on the sunrise and sunset time at 85.446758^o^W. Thus we assign 85.446758 to `lng`.

-   `lat`: Latitude of the study site (in decimal). The provided example focuses on the sunrise and sunset time at 42.476708^o^N. Thus we assign 42.476708 to `lat`.

```{r comment=""}
path  <- "./data/1-1B Switchgrass Time Series.csv"
lng   <- -85.446758
lat   <- 42.476708
```

**Step3:** Read data. Read the CSV file in table format and creates a data frame through the function `read.csv()`. Note that only the first column is used in this example.

```{r comment=""}
dat <- read.csv(path)
head(dat)  #print first parts of the data
```

Then the day, month, and year are extracted from date list through the function `strsplit()`.

```{r comment=""}
day   <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[1]))
                                     # extract day numbers from characters
month <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[2]))
                                     # extract month numbers from characters
year  <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[3]))
                                     # extract year numbers from characters

cat(
  "Dates:",head(dat$date),"...\n",
  "************Extract day, month, and year numbers************\n",
  "Days:",head(day),"...\n",
  "Months:",head(month),"...\n",
  "Years:",head(year),"...\n"
)
```

**Step4:** Run the model. The first step is to calculate day of the year (DOY) using the following equation:

$$
\text{DOY}=\left[ \frac{275\text{month}}{9} \right]-\left[\frac{\text{month}+9}{12}\right]·
\left(\left[ \frac{\text{year}-4\text[\text{year}/4]+2}{3} \right]+1\right)+\text{day}-30,\quad(1)
$$

```{r comment=""}
doys <- DOY(year, month, day)          # calculate day of the year (DOY)

cat(                                   # present the output
  "Dates:", head(dat$date),"...\n",
  "DOYs:", head(doys),"...\n"
)
```

Sunrise time (Time~sr~, hour) is calculated as:

$$
\text{Time}_{sr}=12-\frac{1}{15}\cos^{-1}\left(\frac{-\sin\Phi\sin\delta}{\cos\Phi\cos\delta} \right)-\frac{t}{60},\quad(2)
$$

where $\Phi$ is latitude (decimal), $\delta$ is solar declination (decimal), $t$ is local time (hour) which can be estimated from longitude. Sunset time (Time~ss~, hour) can be estimated by:

$$
\text{Time}_{ss}=12+\frac1{15}\cos^{-1}\left(\frac{-\sin\Phi\sin\delta}{\cos\Phi\cos\delta}\right)-\frac t{60},\quad(3)
$$

Solar declination ($\delta$) is calculated as:

$$
\delta = -23.25·\cos\left[\frac{360}{365}(\text{DOY}+10) \right],\quad(4)
$$

Calculation of Time~sr~ and Time~ss~ via equation (2--4) can be done through the function `sunRS`:

```{r comment=""}
time_sr <- as.numeric(lapply(lapply(doys, sunRS, lng, lat), function(xx) xx[1]))
                                      # calculate sunrise time
time_ss <- as.numeric(lapply(lapply(doys, sunRS, lng, lat), function(xx) xx[2]))
                                      # calculate sunset time

cat(
  paste0("Longitude = ",lng, ", Latitude = ",lat,"\n"),
  "Date:", head(dat$date),"...\n",
  "Sunrise Time:", round(head(time_sr),3),"... (hour)\n",
  "Sunset Time:", round(head(time_ss),3),"... (hour)\n"
)                                     # present the output
```

Now we can combine the calculated results by columns and view them.

```{r comment=""}
res <- data.frame(
  date = dat$date, time_sr, time_ss
)                             # calculated result
                              #  * column 1  ($date): observation date
                              #  * column 2  ($time_sr): sunrise time
                              #  * column 3  ($time_ss): sunset time
print(head(res), digits = 3)
```

**Step5:** Save and visualize results. Use the `write.csv()` function to send outputs to a CSV file.

```{r eval=FALSE}
write.csv(res, 
          "./data/1-4B Switchgrass Sunrise and Sunset Time Output.csv", 
          row.names = FALSE)
```

Create a line plot of calculated variables against time through the function `plot()`:

```{r comment=""}
plot(
  doys, time_sr, col=2, ylim=range(c(time_sr, time_ss)), type="l",
  xlab = "Day of the year", ylab = "Temperature",
  main = "2016 Sunrise and Sunset Time of Kellogg Biological Station"
)                             # plot dew point temperature versus time
lines(doys, time_ss, col=3, type="l")
                              # plot wet-bulb temperature versus time
legend("right", pch=c(15, 15), col=c(2, 3), bty="n",
       legend=c("Sunrise Time", "Sunset Time")) # add legends
```

## Reference

Campbell, G. S., and Norman, J. (2012). An Introduction to Environmental Biophysics. Springer Science & Business Media. 286pp.
