---
title: "Calculating Solar Angles"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate zenith, azimuth, and elevation angles through the function `zangle()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate Solar Angles
# ********************************
#
# Calculate the zenith, azimuth, and elevation angles for a single day. 
#
# ********************************
# List of variables and units for this example
#
# 'year'        experimental year
# ’month'       experimental month
# 'day'         experimental day
# 'hours'       hours of the day (h)
# 'lng'         longitute of the study site (decimal)
# 'lat'         latitude of the study site (decimal)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
year  <- 2016
month <- 9
day   <- 22
hours <- seq(0, 23.5, 0.5)      # 0.5h, 1h, 1.5h, ..., 23.5h
lng   <- 85.446758
lat   <- 42.476708

# Step 3: run the model
res <- zangle(year, month, day, lng, lat, time_series = hours)
                                # calculated result
                                  #  * column 1  ($time): hours of the day
                                  #  * column 2  ($zenith): zenith angle
                                  #  * column 3  ($azimuth): azimuth angle
                                  #  * column 4  ($elevation): elevation angle

# Step 4: show and visualize results
print(res, digits = 3)          # present the results

with(res, plot(
  time, zenith, col=2, ylim=range(c(zenith, azimuth, elevation)), type="l",
  xlab = "Hours of the day", ylab = "Solar Angles"
))                              # plot zenith angle versus time
with(res, lines(time, azimuth, col=3, type="l"))
                                # plot azimuth angle versus time
with(res, lines(time, elevation, col=4, type="l"))
                                # plot elevation angle versus time
legend("right", pch=c(15, 15, 15), col=c(2, 3, 4), bty="n",
       legend=c("Zenith", "Azimuth", "Elevation"))          # add legends

```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `day`, `month`, and `year`: Experimental day. The provided example calculates the solar angles for 22 September 2016 at 42.476708^o^N and, therefore, we assigns 22 to the variable `day`, 9 to `month`, 2016 to `year`.

-   `hours`: Hours of the day used for calculating solar angles. The provided example calculates solar angles in half-hour time resolution. Thus, we assigns time series (0h, 0.5h, 1h, ..., 23.5h) to the variable hours through the function `seq`. Since the solar angles are only valid between the sunrise and sunset time, they will not be estimated during the night-time hours.

-   `lng`: Longitude of the study site (in decimal). The provided example focuses on the solar angles at 85.446758^o^E. Thus we assign 85.446758 to `lng`.

-   `lat`: Latitude of the study site (in decimal). The provided example focuses on the solar angles at 42.476708^o^N. Thus we assign 42.476708 to `lat`.

```{r comment=""}
year  <- 2016
month <- 9
day   <- 22
hours <- seq(0, 23.5, 0.5)      # 0.5h, 1h, 1.5h, ..., 23.5h
lng   <- 85.446758
lat   <- 42.476708
```

**Step3:** Zenith angle ($\psi$, decimal) is calculated from:

$$
\psi=\cos^{-1}(\sin\Phi\sin\delta+\cos\Phi\cos\delta\cos(15(t-t_0))),\quad(2)
$$

where $\Phi$ is latitude (decimal), $\delta$ is solar declination, $t$ is local time that can be estimated from longitude, $t_0$ is the time of solar noon. Solar declination $\delta$ can be estimated by:

$$
\delta=-23.25·\cos\left[ \frac{360}{365}(\text{DOY}+10)\right],\quad(3)
$$

Azimuth angle ($A_z$, decimal) can be calculated as:

$$
A_z=\cos^{-1}\left( \frac{-\sin\delta-\cos\delta\sin\psi\sin\Phi}{\cos\Phi\sin\psi} \right),\quad(4)
$$

Solar elevation angle can be calculated as the difference between 90^o^ and the zenith angle.

```{r comment=""}
res <- zangle(year, month, day, lng, lat, time_series = hours)
                                # calculated result
                                  #  * column 1  ($time): hours of the day
                                  #  * column 2  ($zenith): zenith angle
                                  #  * column 3  ($azimuth): azimuth angle
                                  #  * column 4  ($elevation): elevation angle

print(res, digits = 3)
```

**Step4:** Visualize the result. Create a line plot of predicted solar angles against time through the function `plot()`:

```{r comment=""}
with(res, plot(
  time, zenith, col=2, ylim=range(c(zenith, azimuth, elevation)), type="l",
  xlab = "Hours of the day", ylab = "Solar Angles"
))                              # plot zenith angle versus time
with(res, lines(time, azimuth, col=3, type="l"))
                                # plot azimuth angle versus time
with(res, lines(time, elevation, col=4, type="l"))
                                # plot elevation angle versus time
legend("right", pch=c(15, 15, 15), col=c(2, 3, 4), bty="n",
       legend=c("Zenith", "Azimuth", "Elevation"))          # add legends
```
