---
title: "Calculating Solar Angles of a Time Period"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate zenith, azimuth, and elevation angles of a period through the function `zangle()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate Solar Angles at noon for 2016 at Kellogg Biological Station
# ********************************
#
# Calculate the zenith, azimuth, and elevation angles at noon during the course of
# year 2016 at Kellogg Biological Station (KBS), Michigan, USA.
#
# ********************************
# List of variables and units for this example
#
# 'path'        path of a CSV file, which conains one column (with table titles)
#                * column 1: observation date. Required format "dd/mm/yyyy"
# 'hours'       hours of the day (h)
# 'lng'         longitute of the study site (decimal)
# 'lat'         latitude of the study site (decimal)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
path  <- "./data/1-1B Switchgrass Time Series.csv"
hour  <- 12                     # noon
lng   <- -85.446758
lat   <- 42.476708

# Step 3: read data
dat   <- read.csv(path)              # read the CSV file
                                     #  * column 1 ($date): observation date
                                     #  * column 2 ($t_max): not used
                                     #  * column 3 ($t_min): not used
day   <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[1]))
                                     # extract day numbers from characters
month <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[2]))
                                     # extract month numbers from characters
year  <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[3]))
                                     # extract year numbers from characters

# Step 4: run the model
res <- do.call(rbind, lapply(1:nrow(dat), function(i){
  data.frame(
    date = dat$date[i],
    zangle(year[i], month[i], day[i], lng, lat, time_series = hour)
  )
}))                             # calculated result
                                  #  * column 1  ($date): observation date
                                  #  * column 2  ($time): hours of the day
                                  #  * column 3  ($zenith): zenith angle
                                  #  * column 4  ($azimuth): azimuth angle
                                  #  * column 5  ($elevation): elevation angle

# Step 5: show and visualize results
print(res, digits = 3)          # present the results

with(res, plot(
  seq_len(nrow(dat)), zenith, col=2, ylim=range(c(zenith, azimuth, elevation)), type="l",
  xlab = "Day of the year", ylab = "Solar Angles"
))                              # plot zenith angle versus time
with(res, lines(seq_len(nrow(dat)), azimuth, col=3, type="l"))
                                # plot azimuth angle versus time
with(res, lines(seq_len(nrow(dat)), elevation, col=4, type="l"))
                                # plot elevation angle versus time
legend("right", pch=c(15, 15, 15), col=c(2, 3, 4), bty="n",
       legend=c("Zenith", "Azimuth", "Elevation"))          # add legends
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `path`: the path of a CSV file which the data are to be read from. Each row of the file appears as one line, and columns are separated by commas (,). The first row appears as the table title. There should be three columns in this file:

    -   Column 1: Observation date. The format should be 'dd/mm/yyyy', where 'dd' refers to day, 'mm' refers to month and 'yyyy' refers to the year number. Even though the sample file "1-1B Switchgrass Time Series.csv" contains three columns, the last two columns are not used in this example.

-   `hours`: Hours of the day used for calculating solar angles. The provided example calculates solar angles at noon. Thus, we assigns 12h to `hours`.

-   `lng`: Longitude of the study site (in decimal). The provided example focuses on the solar angles at 85.446758^o^E. Thus we assign 85.446758 to `lng`.

-   `lat`: Latitude of the study site (in decimal). The provided example focuses on the solar angles at 42.476708^o^N. Thus we assign 42.476708 to `lat`.

```{r comment=""}
path  <- "./data/1-1B Switchgrass Time Series.csv"
hour <- 12
lng   <- -85.446758
lat   <- 42.476708
```

**Step3:** Read data. Read the CSV file in table format and creates a data frame through the function `read.csv()`. Note that only the first column is used in this example.

```{r comment=""}
dat <- read.csv(path)
head(dat)  #print first parts of the data
```

Then the day, month, and year are extracted from date list through the function `strsplit()`.

```{r comment=""}
day   <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[1]))
                                     # extract day numbers from characters
month <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[2]))
                                     # extract month numbers from characters
year  <- as.numeric(lapply(strsplit(dat$date,"/"),function(xx) xx[3]))
                                     # extract year numbers from characters

cat(
  "Dates:",head(dat$date),"...\n",
  "************Extract day, month, and year numbers************\n",
  "Days:",head(day),"...\n",
  "Months:",head(month),"...\n",
  "Years:",head(year),"...\n"
)
```

**Step4:** Zenith angle ($\psi$, decimal) is calculated from:

$$
\psi=\cos^{-1}(\sin\Phi\sin\delta+\cos\Phi\cos\delta\cos(15(t-t_0))),\quad(2)
$$

where $\Phi$ is latitude (decimal), $\delta$ is solar declination, $t$ is local time that can be estimated from longitude, $t_0$ is the time of solar noon. Solar declination $\delta$ can be estimated by:

$$
\delta=-23.25·\cos\left[ \frac{360}{365}(\text{DOY}+10)\right],\quad(3)
$$

Azimuth angle ($A_z$, decimal) can be calculated as:

$$
A_z=\cos^{-1}\left( \frac{-\sin\delta-\cos\delta\sin\psi\sin\Phi}{\cos\Phi\sin\psi} \right),\quad(4)
$$

Solar elevation angle can be calculated as the difference between 90^o^ and the zenith angle.

```{r comment=""}
res <- do.call(rbind, lapply(1:nrow(dat), function(i){
  data.frame(
    date = dat$date[i],
    zangle(year[i], month[i], day[i], lng, lat, time_series = hour)
  )
}))                             # calculated result
                                  #  * column 1  ($date): observation date
                                  #  * column 2  ($time): hours of the day
                                  #  * column 3  ($zenith): zenith angle
                                  #  * column 4  ($azimuth): azimuth angle
                                  #  * column 5  ($elevation): elevation angle

print(head(res), digits = 3)
```

**Step5:** Save and visualize results. Use the `write.csv()` function to send outputs to a CSV file.

```{r eval=FALSE}
write.csv(res, 
          "./data/1-5B Switchgrass Solar Angles Output.csv", 
          row.names = FALSE)
```

Create a line plot of predicted solar angles against time through the function `plot()`:

```{r comment=""}
with(res, plot(
  seq_len(nrow(dat)), zenith, col=2, ylim=range(c(zenith, azimuth, elevation)), type="l",
  xlab = "Day of the year", ylab = "Solar Angles"
))                              # plot zenith angle versus time
with(res, lines(seq_len(nrow(dat)), azimuth, col=3, type="l"))
                                # plot azimuth angle versus time
with(res, lines(seq_len(nrow(dat)), elevation, col=4, type="l"))
                                # plot elevation angle versus time
legend("right", pch=c(15, 15, 15), col=c(2, 3, 4), bty="n",
       legend=c("Zenith", "Azimuth", "Elevation"))          # add legends
```
