---
title: "Heat Flux and Heat Storages"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate heat flux density through the function `heatflux()` and heat storage through the function `heatstore()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate Heat Flux Density and Heat Storeage
# ********************************
#
# Calculate heat flux density and heat storage based on their definitions.
#
# ********************************
# List of variables and units for this example
#
# * Variables for calculating heat flux density
#
# 'dt_sd'       soil temperature change when heat passes through a thin layer
#                 (Celsius, Fahrenheit, or Kelvin)
# ’d'           thickness of the soil layer (m)
#
# * Variables for calculating heat storage
#
# 'dt_st'       soil temperature change during a time interval
#                 (Celsius, Fahrenheit, or Kelvin)
# 'dsecond'     time interval (s)
# 'theta'       soil water content (%)
#
# Calculated variables in the program
# 'G'           heat flux density (W m−2)
# 'dS'          heat storage (W m−3)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
dt_sd <- 2      # soil temperature change (for calculating heat flux density)
d <- 1          # thickness of the soil layer (for calculating heat flux density)

dt_st <- 2      # soil temperature change (for calculating heat storage)
dsecond <- 3600 # time inteval (for calculating heat storage)
theta <- 50     # soil water content (for calculating heat storage)

# Step 3: run the model
G <- heatflux(dt_sd, d)
dS <- heatstore(dt_st, dsecond, theta)

# Step 4: present the output
cat(
  paste("***** Calculating heat flux density *****\n"),
  paste("Soil temperature change ΔT =", dt_sd, "(Celsius)\n"),
  paste("Thickness of the soil laryer d =", d, "(m)\n"),
  paste("Heat flux density G = 2.5 * ΔT / d =", G, "(W m-2)\n"),
  paste("\n"),
  paste("***** Calculating heat storage *****\n"),
  paste("Soil temperature change ΔT =", dt_st, "(Celsius)\n"),
  paste("Time inteval = Δt", dsecond, "(s)\n"),
  paste("Soil water content = θ", theta, "(%)\n"),
  paste("Heat storage ΔS = (1000 * 890 + theta / 100 * 1000 * 4190) * (ΔT / Δt) =", round(dS, 2), "(W m-3)\n")
)

```

## Explanations

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters. For calculating heat flux density, the following parameters are required:

-   `dt_sd`: The temperature difference between the surface and bottom soil layer (Celsius, Fahrenheit, or Kelvin). The observed value is 2^o^C in this example.

-   `d`: Thickness of the soil layer (m). In this example it is 2m.

To calculate the amount of heat energy stored in the soil, we should specify:

-   `dt_st`: The change on temperature over a period of time (Celsius, Fahrenheit, or Kelvin). The observed value is 2^o^C in this example.

-   `dsecond`: Time interval (s). The observation lasts for 1h (=3600s) in this example.

-   `theta`: Soil water content (%). The provided example uses 50%.

```{r comment=""}
dt_sd <- 2      # soil temperature change (for calculating heat flux density)
d <- 1          # thickness of the soil layer (for calculating heat flux density)

dt_st <- 2      # soil temperature change (for calculating heat storage)
dsecond <- 3600 # time inteval (for calculating heat storage)
theta <- 50     # soil water content (for calculating heat storage)
```

**Step3:** Run the model. Heat flux density ($G$, W m^−2^ ) is calculated as:

$$G=\kappa·\frac{\Delta T}{d},\quad(1)$$

where $\kappa$ (W m^−1^ K^−1^) is the thermal conductivity of the soil and $d$ (m) is the thickness of the soil layer. The $kappa$ value is ∼ 2.5 W m^−1^ K^−1^ for soil minerals, ∼1.92 W m^−1^ K^−1^ for organic matter, and 4.18 W m^−1^ K^−1^ for water.

```{r}
G <- heatflux(dt_sd, d, kappa = 2.5)

cat(
  paste("***** Calculating heat flux density *****\n"),
  paste("Soil temperature change ΔT =", dt_sd, "(Celsius)\n"),
  paste("Thickness of the soil laryer d =", d, "(m)\n"),
  paste("Heat flux density G = 2.5 * ΔT / d =", G, "(W m-2)\n")
)
```

The arguments and their meaning of the function `heatflux()` is shown in the table below:

| Argument | Data Type | Meaning                                                                                        | Symbol     |
|---------------|---------------|---------------------------|---------------|
| `detT`   | vector    | soil temperature change when heat passes through a thin layer (Celsius, Fahrenheit, or Kelvin) | $\Delta T$ |
| `d`      | vector    | thickness of the soil layer (m)                                                                | $d$        |
| `kappa`  | scalar    | the thermal conductivity constant                                                              | $\kappa$   |

The heat storage ($\Delta S$, W m^−3^) over a period of time ($\Delta t$, s) can be calculated as:

$$\Delta S=(ρ_b·c_d+\theta·ρ_w·c_w)·\frac{\Delta T}{\Delta t}·d$$ where $\rho_b$ (kg m^-1^) is the soil bulk density (default 2.7 kg m^-3^), $\rho_w$ is the density of water, $c_d$ (890 J kg^−1^ K^−1^) and $c_w$ (4190 J kg^−1^ K^−1^) are the specific heat capacities of the dry mineral soil and the soil water, respectively. $\theta$ is the volumetric soil water content (%), and $∆T/∆t$ (K s^−1^) is the mean soil temperature change at the time interval.

```{r}
dS <- heatstore(dt_st, dsecond, theta)

cat(
  paste("***** Calculating heat storage *****\n"),
  paste("Soil temperature change ΔT =", dt_st, "(Celsius)\n"),
  paste("Time inteval = Δt", dsecond, "(s)\n"),
  paste("Soil water content = θ", theta, "(%)\n"),
  paste("Heat storage ΔS = (1000 * 890 + theta / 100 * 1000 * 4190) * (ΔT / Δt) =", round(dS, 2), "(W m-3)\n")
)
```

The arguments and their meaning of the function `heatstore()` are shown in the table below:

| Argument  | Data Type | Meaning                                                                         | Symbol     |
|---------------|---------------|---------------------------|---------------|
| `detT`    | vector    | soil temperature change during a time interval (Celsius, Fahrenheit, or Kelvin) | $\Delta T$ |
| `dettime` | vector    | time interval (s)                                                               | $\Delta t$ |
| `theta`   | vector    | soil water content (%)                                                          | $\theta$   |
| `roub`    | scalar    | soil bulk density (kg m^-3^). Default 2.7.                                      | $\rho_b$   |
