---
title: "Calculate Wind Speed at the Standard Height"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate wind speed at the standard height (2 meters) through the function `ws()`.

## Code Example(s)

```{r eval=FALSE}
# Calculate Wind Speed at the Standard Height
# ********************************
#
# Calculate wind speed at 2 meters from wind speed measured at a certain height.
#
# ********************************
# List of variables and units for this example
#
# 'z'         height above ground level (m)
# 'uz'        horizontal wind speed (m s−1) at height z (m)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
z <- 9
uz <- 1.60

# Step 3: run the model
u2 <- ws(z, uz)

# Step 4: present the output
cat(
  paste("Measured wind speed at", z, "m:", uz, "(m s-1)\n"),
  paste("Wind speed at 2 meters:", round(u2, 2), "(m s-1)\n")
)
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `z`: Height above ground level (m). In this example, the wind speed is measured at 9m, so we assign 9 to `z`.
-   `uz`: Wind speed measured at a certain height (m s^-1^). The measured value of wind speed at 9m is 1.60.

```{r comment=""}
z <- 9
uz <- 1.60
```

**Step3:** Run the model. As discussed in Section 1-7A, the wind profile of homogeneous vegetation under neutral atmospheric conditions is typically expressed as:

$$U(z)=\frac{u^*}{\kappa}·\ln(\frac{z-d}{z_0}),\quad(1)$$

where $U(z)$ is the horizontal wind speed (m s^−1^) at height $z$ (m); $u^∗$ is the friction velocity (m s^−1^ ), $\kappa$ is the von Karmon constant with an average value of 0.35 -- 0.43 (a value of 0.40 is often used in the literature), $d$ (m) is a zero plain displacement, and $z_0$ is the roughness length (m) at which is $U$ is near zero. Thus, the wind speed at 2 meters can be calculated as:

$$
U(2)=U(z)\frac{\ln ((2-d)/z_0)}{\ln ((z-d)/z_0)},\quad(2)
$$

```{r comment=""}
u2 <- ws(z, uz)

cat(
  paste("Measured wind speed at", z, "m:", uz, "(m s-1)\n"),
  paste("Wind speed at 2 meters:", round(u2, 2), "(m s-1)\n")
)
```

The arguments and their meaning of the function `ws()` that might change the calculated result in this example are shown in the table below:

| Argument | Data Type | Meaning                                                                                                                      | Symbol |
|----------------|----------------|------------------------|----------------|
| `z`      | vector    | height above ground level (m)                                                                                                | $z$    |
| `uz`     | vector    | wind speed measured at a certain height (m s^-1^)                                                                            | $U(z)$ |
| `d`      | vector    | zero plain displacement (m). Default 0.08                                                                                    | $d$    |
| `z0`     | vector    | roughness length (m) at which wind speed is near zero. Default 0.01                                                          | $z_0$  |
| `mode`   | character | options of analysis. Must be set as `"wind2"` to generate wind speed profile. This is the defaulted value of argument `mode` | /      |
