---
title: "Landsberg Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate ecosystem production based on Landsberg model.

## Code Example(s)

```{r eval=FALSE}
# Ecosystem Production based on Landsberg Model
# *****************************************************************
#
# Calculate ecosystem production based on Michaelis-Menten (MM) Model (Landsberg,
# 1977).
#
# ********************************
# List of variables and units for this example
#
# 'PAR'        	  photosynthetically active radiation (PAR, μmol m-2 s-1)
# 'alpha'         slope of the change in photosynthesis rate with PAR
# 'Pm'            maximum photosynthetic capacity (μmol m-2 s-1)
# 'Icomp'         light compensation point (μmol m-2 s-1)
#
# Calculated variables in the program
# 'Pn'            photosynthesis rate (μmol m-2 s-1) given by Landsberg model
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
PAR   <- seq(0, 1000, 10)
alpha <- 0.008
Pm    <- 10
Icomp <- 200

# Step 3: run the model
Pn <- photo.landsberg(PAR, alpha = alpha, Pm = Pm, Icomp = Icomp)

# Step 4: present and visualize the result
print(data.frame(PAR = PAR, Pn = Pn), digits = 3)

plot(
  PAR, Pn, type="l",
  xlab = "Photosynthetically active radiation",
  ylab = "Photosynthesis rate",
  main = "Photosynthesis Rate Estimated by Landsberg Model"
)                       # plot photosynthesis rate versus PAR (MM model)

```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `PAR`: Photosynthetically active radiation ($PAR$, µmol m^−2^ s^−1^). The provided example calculates photosynthesis rate ($P_n$, µmol m^−2^ s^−1^) when $PAR$ changes from 0 to 1000 µmol m^−2^ s^−1^.

-   `alpha`: Slope of the change in photosynthesis rate with $PAR$. The provided example uses 0.008.

-   `Pm`: Maximum photosynthetic capacity (µmol m^−2^ s^−1^). The provided example uses 10 µmol m^−2^ s^−1^.

-   `Icomp`: Light compensation point (μmol m^−2^ s^−1^) at which the photosynthesis rate is zero. The provided example uses 200.

```{r comment=""}
PAR   <- seq(0, 1000, 10)
alpha <- 0.008
Pm    <- 10
Icomp <- 200
```

**Step3:** Run the model. The Landsberg model is expressed as (Landsberg, 1977):

$$
P_n=P_m·[1-e^{\alpha·(PAR-I_{comp})}]
$$

where $P_m$ (µmol m^−2^ s^−1^) is the maximum photosynthesis, *α* is the slope of the change in $P_n$ (µmol m^−2^ s^−1^) with $PAR$ (i.e., a shape factor), and $I_{comp}$ (µmol m^−2^ s^−1^) is the light compensation point at which $P_n$ is zero. The following example shows how to calculate $P_n$ using the Landsberg model:

```{r comment=""}
Pn <- photo.landsberg(PAR, alpha = alpha, Pm = Pm, Icomp = Icomp)

print(head(data.frame(PAR = PAR, Pn = Pn)), digits = 3)
```

**Step4:** Visualize results. Create a line plot of $P_n$ against $PAR$ through the function `plot()`:

```{r comment=""}
plot(
  PAR, Pn, type="l",
  xlab = "Photosynthetically active radiation",
  ylab = "Photosynthesis rate",
  main = "Photosynthesis Rate Estimated by Landsberg Model"
)                       # plot photosynthesis rate versus PAR (MM model)
```

## Reference

Landsberg, J. J. (1977). Some useful equations for biological studies. Experimental Agriculture, 13(3), 273--286.
