---
title: "Fit Time Series Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to fit time series model through the function `respir.fit()`.

## Code Example(s)

```{r eval=FALSE}
# Fit Time Series Model
# ********************************
#
# Fit time series model from field measurements of soil temperature at 5cm, soil
# moisture at 10 cm, day of the year, and soil respiration, at a larch forest in 
# Mt. Fuji in 2015. Data provided by Dr. Naishen Liang (Email: liang@nies.go.jp)
#
# ********************************
# List of variables and units for this example
#
# 'path'        path of a CSV file, which contains four columns (with table titles)
#                * column 1(doy):    day of the year
#                * column 2(t_s):    soil temperature (Celsius)
#                * column 3(theta):  soil water content (%). Valid from 0-100
#                * column 4(Rs):     ecosystem respiration (µmol CO2 m−2 s−1)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
path  <- "./data/3-1G Simulated Soil Respiration.csv"

# Step 3: read data
dat   <- read.csv(path) # read the CSV file
                  #  * column 1 ($doy): day of the year
                  #  * column 2 ($t_s): temperature (Celsius)
                  #  * column 3 ($theta): soil water content (%). Valid from 0-100
                  #  * column 4 ($Rs): ecosystem respiration (µmol CO2 m−2 s−1)

# Step 4: fit the model
md <- respir.fit(dat$Rs, dat$t_s, dat$theta, dat$doy, method = "tmd.Xu")
                  #  Model 1 (Xu et al., 2011)

cat(              # summarize the model fit
  paste("***** Xu et al. (2011) *****\n"),
  paste("alpha * beta1 =", round(md$coef[["alpha_plus_beta1"]], 3), "\n"),
  paste("beta0 =", round(md$coef[["beta0"]], 3), "\n"),
  paste("beta2 =", round(md$coef[["beta2"]], 3), "\n"),
  paste("beta3 =", round(md$coef[["beta3"]], 3), "\n"),
  paste("beta4 =", round(md$coef[["beta4"]], 3), "\n"),
  paste("Multiple R-squared:", round(md$r.squared, 3), "\n"),
  paste("Adjusted R-squared:", round(md$adj.r.squared, 3), "\n")
)

# Step 5: model prediction
ta  <- dat$t_s                    # soil temperature
m   <- dat$theta                  # soil water content
doy <- dat$doy                    # day of the year

alpha <- md$coef[["alpha_plus_beta1"]]
beta1 <- 1
beta0 <- md$coef[["beta0"]]      # extract model parameter
beta2 <- md$coef[["beta2"]]      # extract model parameter
beta3 <- md$coef[["beta3"]]      # extract model parameter
beta4 <- md$coef[["beta4"]]      # extract model parameter
eval(parse(text = md$formula))   # make predictions
                                 # results are stored in the variable 'R'

plot(dat$Rs - R, pch = 19, cex=0.5, xaxt = "n",
     xlab = "Observation", ylab = "Residue",
     main = "Residue Plot for Xu's Model")  # Residue plot
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `path`: the path of a CSV file which the data are to be read from. Each row of the file appears as one line, and columns are separated by commas (,). The first row appears as the table title. The sample file "3-1 Soil Respiration Measurements in Mt. Fuji.csv" contains four columns:

    -   Column 1 (`$doy`): Day of the year.

    -   Column 2 (`$t_s`): Soil temperature measured at 5cm (Celsius).

    -   Column 3 (`$theta`): Soil water content (%). Valid from 0--100.

    -   Column (`$Rs`): Measured soil respiration (µmol CO~2~ m^−2^ s^−1^).

```{r comment=""}
path  <- "./data/3-1G Simulated Soil Respiration.csv"
```

**Step3:** Read data. Read the CSV file in table format and creates a data frame through the function `read.csv()`.

```{r comment=""}
dat   <- read.csv(path) # read the CSV file
                  #  * column 1 ($doy): day of the year
                  #  * column 2 ($t_s): temperature (Celsius)
                  #  * column 3 ($theta): soil water content (%). Valid from 0-100
                  #  * column 4 ($Rs): ecosystem respiration (µmol CO2 m−2 s−1)

head(dat)         # print first parts of the data
```

**Step4:** Fit the model. Soil respiration ($R_s$, µmol CO~2~ m^−2^ s^−1^) can be modeled as a function of soil temperature ($T$, ^o^C), soil water content ($\theta$, 0--100%), and day of the year (DOY) simultaneously (Xu et al., 2011):

$$
R=\alpha·e^{\beta_0·T}+\beta_1·(\theta-\beta_2)^2+\beta_3·(DOY-\beta_4)^2,\quad(1)
$$

```{r comment=""}
md <- respir.fit(dat$Rs, dat$t_s, dat$theta, dat$doy, method = "tmd.Xu")
```

Now summarize the model fit:

```{r comment=""}
cat(              # summarize the model fit
  paste("***** Xu et al. (2011) *****\n"),
  paste("alpha * beta1 =", round(md$coef[["alpha_plus_beta1"]], 3), "\n"),
  paste("beta0 =", round(md$coef[["beta0"]], 3), "\n"),
  paste("beta2 =", round(md$coef[["beta2"]], 3), "\n"),
  paste("beta3 =", round(md$coef[["beta3"]], 3), "\n"),
  paste("beta4 =", round(md$coef[["beta4"]], 3), "\n"),
  paste("Multiple R-squared:", round(md$r.squared, 3), "\n"),
  paste("Adjusted R-squared:", round(md$adj.r.squared, 3), "\n")
)
```

**Step5:** Model prediction. Predict $R_s$ using the fitted model:

```{r comment=""}
ta  <- dat$t_s                    # soil temperature
m   <- dat$theta                  # soil water content
doy <- dat$doy                    # day of the year

alpha <- md$coef[["alpha_plus_beta1"]]
beta1 <- 1
beta0 <- md$coef[["beta0"]]      # extract model parameter
beta2 <- md$coef[["beta2"]]      # extract model parameter
beta3 <- md$coef[["beta3"]]      # extract model parameter
beta4 <- md$coef[["beta4"]]      # extract model parameter
eval(parse(text = md$formula))   # make predictions
                                 # results are stored in the variable 'R'

print(head(data.frame(
  t_s = dat$t_s, theta = dat$theta, doy = dat$doy,
  predicted_Rs = R
)), digits = 3)
```

Create a residue plot of fitted model:

```{r comment=""}
plot(dat$Rs - R, pch = 19, cex=0.5, xaxt = "n",
     xlab = "Observation", ylab = "Residue",
     main = "Residue Plot for Xu's Model")  # Residue plot
```

The residue is defined as the difference between the observed and predicted value. Positive values means that the model overestimates the soil respiration, whilst negative values indicate underestimation. It can be seen that the residues given by Xu's model are pretty symmetrically distributed towards the line at 0, which indicates that this model gives a good fit for the given data set.

## Reference

Xu, J., Chen, J., Brosofske, K., Li, Q., Weintraub, M., Henderson, R., Wilske, B., John, R., Jensen, R., Li, H., and Shao, C. (2011). Influence of timber harvesting alternatives on forest soil respiration and its biophysical regulatory factors over a 5-year period in the Missouri Ozarks. Ecosystems, 14(8), 1310--1327.

