---
title: "Calculate Potential Evapotranspiration Using Thornthwaite Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate potential evapotranspiration (PET) using Thornthwaite model.

```{r eval=FALSE}
# Calculate PET Using Thornthwaite Model
# ********************************
#
# Calculate PET based on Thornthwaite Model (Thornthwaite, 1948).
#
# ********************************
# List of variables and units for this example
#
# 't_a'      monthly mean air temperature (Celsius)
# 'Ld'       monthly mean day length (h)
# 'PET'      monthly potential evapotranspiration (cm)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat   <- read.csv("./data/4-1B Simulated Monthly Micrometeorological Varibles.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Month):   month of the year
              #     Column 2 ($Ta):      monthly mean air temperature (Celsius)
              #     Column 3 ($Ld):      monthly mean day length (h)

t_a  <- dat$Ta         # monthly mean air temperature
Ld   <- dat$Ld/12      # monthly mean daytime length

# Step 3: run the model
PET <- m2.PET.Thornthwaite(t_a, Ld)

# Step 4: present and visualize results
print(head(data.frame(
  t_a, Ld, PET
)), digits = 3)     # summarize the result

plot(
  1:12, PET, type = "l", 
  xlab = "Month of the year", 
  ylab = "Monthly PET",
  main = "PET Estimated by Thornthwaite Model"
)                   # plot PET versus time
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv`:

```{r comment=""}
dat   <- read.csv("./data/4-1B Simulated Monthly Micrometeorological Varibles.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Month):   month of the year
              #     Column 2 ($Ta):      monthly mean air temperature (Celsius)
              #     Column 3 ($Ld):      monthly mean day length (h)

head(dat)
```

Extract micrometeorological variables that are required in Thornthwaite Model from variable `dat`.

```{r comment=""}
t_a  <- dat$Ta         # monthly mean air temperature
Ld   <- dat$Ld/12      # monthly mean daytime length
```

Note that factor 12 is required for unit conversion from multiples of 12h to hour.

**Step3:** Run the model. The Thornthwaite (1948) PET model is expressed as:

$$
\text{PET}=1.6·L_d·\left( \frac{10·T}{I}\right)^a,\quad(1)
$$

where:

-   $\text{PET}$ = monthly PET (cm)

-   $L_d$ = monthly mean daytime length in multiples of 12h

-   $a = 6.75 × 10{-7}·I^3−7.71×10^{−5}·I^2+0.01792·I+0.49239$

-   $I$ = annual heat index, which is computed from the monthly heat indices

$$
I = \sum_{j=1}^{12}i_j,\quad(2)
$$

where $i_j=\left(\frac{T_j}5\right)^{1.514}$ , $T_j$ is the mean air temperature (^o^C) per month.

```{r comment=""}
PET <- m2.PET.Thornthwaite(t_a, Ld)

print(head(data.frame(
  t_a, Ld, PET
)), digits = 3)     # summarize the result
```

**Step4:** Visualize results. Create a line plot of predicted PET against time through the function `plot()`:

```{r comment=""}
plot(
  1:12, PET, type = "l", 
  xlab = "Month of the year", 
  ylab = "Monthly PET",
  main = "PET Estimated by Thornthwaite Model"
)                   # plot PET versus time
```

## References

Thornthwaite, C. W. (1948). An approach toward a Rational classification of climate. Geographical Review, 38(1), 55--94.

Zenone, T., Chen, J., Deal, M. W., Wilske, B., Jasrotia, P., Xu, J., Bhardwaj, A. K., Hamilton, S. K., and Robertson, G. P. (2011). CO~2~ fluxes of transitional bioenergy crops: Effect of land conversion during the first year of cultivation. Global Change Biology-Bioenergy, 3(5), 401--412.
