---
title: "Calculate Potential Evapotranspiration Using Blaney-Criddle Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate potential evapotranspiration (PET) using Blaney-Criddle model.

```{r eval=FALSE}
# Calculate PET Using Blaney-Criddle Model
# ********************************
#
# Calculate PET based on Blaney-Criddle Model (Blaney and Criddle, 1957). Data is
# from field measurements of micrometeorological variables in 2016 in an
# agricultural site with an eddy-covariance tower of the Kellogg Biological Station
# , Michigan, USA (Zenone et al. 2011).
#
# ********************************
# List of variables and units for this example
#
# 't_a'      daily mean air temperature (Celsius)
# 'PET'      daily potential evapotranspiration (mm)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     day of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

t_a <- dat$Ta       # daily mean air temperature

# Step 3: run the model
PET <- m4.PET.BlaneyCriddle(t_a, dhour = 12)

# Step 4: present and visualize results
print(head(data.frame(
  t_a,  PET
)), digits = 3)     # summarize the result

plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Blaney-Criddle Model"
)                   # plot PET versus time
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv`.

```{r comment=""}
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     month of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

head(dat)
```

Extract micrometeorological variables that are required in Blaney-Criddle model from variable `dat`.

```{r comment=""}
t_a <- dat$Ta       # daily mean air temperature
```

**Step3:** Run the model. The Blaney-Criddle equation has the following form (Blaney and Criddle, 1957):

$$
\text{PET}=P·(0.46·T+8.13),\quad(1)
$$

where:

-   $\text{PET}$ = potential water use for a reference crop for a time period (mm)

-   $T$ = mean temperature of a time period (^o^C)

-   $P$ = the percentage of total daytime hours for the period used (daily or monthly) out of total daytime hours of the year (365×12 = 4380 h)

Function `m4.PET.BlaneyCriddle` has two arguments. The second argument specifies the total daytime ours for the period used. When calculating daily PET, we used 12.

```{r comment=""}
PET <- m4.PET.BlaneyCriddle(t_a, dhour = 12)

# Step 4: present and visualize results
print(head(data.frame(
  t_a,  PET
)), digits = 3)     # summarize the result
```

**Step4:** Visualize results. Create a line plot of predicted PET against time through the function `plot()`:

```{r comment=""}
plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Blaney-Criddle Model"
)                   # plot PET versus time
```

## References

Blaney, H. F., and Criddle, W. D. (1957). Report on irrigation water requirements for Pakistan. Bull IDFCR Council, 3(2).

Zenone, T., Chen, J., Deal, M. W., Wilske, B., Jasrotia, P., Xu, J., Bhardwaj, A. K., Hamilton, S. K., and Robertson, G. P. (2011). CO2 fluxes of transitional bioenergy crops: Effect of land conversion during the first year of cultivation. Global Change Biology-Bioenergy, 3(5), 401--412.
