---
title: "Calculate Potential Evapotranspiration Using Turc Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate potential evapotranspiration (PET) using Turc model.

```{r eval=FALSE}
# Calculate PET Using Turc Model
# ********************************
#
# Calculate PET based on Turc Model (Turc, 1961). Data is from field
# measurements of micrometeorological variables in 2016 in an agricultural site
# with an eddy-covariance tower of the Kellogg Biological Station, Michigan, USA
# (Zenone et al. 2011).
#
# ********************************
# List of variables and units for this example
#
# 't_a'      daily mean air temperature (Celsius)
# 'h'        relative humidity in percentage (%)
# 'Rs'       incoming solar radiation (MJ m-2 d-1)
# 'PET'      daily potential evapotranspiration (mm)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     day of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

t_a <- dat$Ta       # daily mean air temperature
h   <- dat$h * 100  # relative humidity in percentage (%)
Rs  <- dat$Rn * 24 * 3600/1e6
                    # incoming solar radiation (convert W m-2 to MJ m-2 d-1)
                    #    NOTE: use net solar radiation as a proxy

# Step 3: run the model
PET <- m5.PET.Turc(t_a, h, Rs)

# Step 4: present and visualize results
print(head(data.frame(
  t_a, h, Rs, PET
)), digits = 3)     # summarize the result

plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Turc Model"
)                   # plot PET versus time
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv` and specify longitude and latitude of the study site.

```{r comment=""}
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     day of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

head(dat)
```

Extract micrometeorological variables that are required in Turc model from variable `dat`.

```{r comment=""}
t_a <- dat$Ta       # daily mean air temperature
h   <- dat$h * 100  # relative humidity in percentage (%)
Rs  <- dat$Rn * 24 * 3600/1e6
                    # incoming solar radiation (convert W m-2 to MJ m-2 d-1)
                    #    NOTE: use net solar radiation as a proxy
```

**Step3:** Run the model. When relative humidity ($h$) is \< 50%, Turc PET model (Turc, 1961) is express as:

$$
\text{PET}=0.013\left(\frac{T}{T+15}\right)(R_s+50)\left(1+\frac{50-h}{70}\right),\quad(1)
$$

when $h$ \> 50%,

$$
\text{PET}=0.013\left(\frac{T}{T+15}\right)(R_s+50),\quad(2)
$$

where:

-   $T$ = daily mean air temperature (^o^C)

-   $R_s$ = daily solar radiation (ly d^−1^, or cal cm^−2^ d^−1^). However, the unit will automatically be converted into MJ m^−2^ d^−1^ through the function `m5.PET.Turc`

-   $h$ = daily mean relative humidity in percentage (%)

```{r comment=""}
PET <- m5.PET.Turc(t_a, h, Rs)

print(head(data.frame(
  t_a, h, Rs, PET
)), digits = 3)     # summarize the result
```

**Step4:** Visualize results. Create a line plot of predicted PET against time through the function `plot()`:

```{r comment=""}
plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Turc Model"
)                   # plot PET versus time
```

## References

Turc, L. (1961). Estimation of irrigation water requirements, potential evapotranspiration: A simple climatic formula evolved up to date. Annuals of Agronomy, 12(1), 13--49.

Jasrotia, P., Xu, J., Bhardwaj, A. K., Hamilton, S. K., and Robertson, G. P. (2011). CO2 fluxes of transitional bioenergy crops: Effect of land conversion during the first year of cultivation. Global Change Biology-Bioenergy, 3(5), 401--412.
