---
title: "Calculate Potential Evapotranspiration Using Priestley-Taylor Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate potential evapotranspiration (PET) using Priestley-Taylor model.

```{r eval=FALSE}
# Calculate PET Using Priestley-Taylor Model
# ********************************
#
# Calculate PET based on Priestley-Taylor Model  (Priestley and Taylor， 1972).
# Data is from field measurements of micrometeorological variables in 2016 in an
# agricultural site with an eddy-covariance tower of the Kellogg Biological
# Station, Michigan, USA (Zenone et al. 2011).
#
# ********************************
# List of variables and units for this example
#
# 't_a'      daily mean air temperature (Celsius)
# 'Rn'       net solar radiation (W m-2)
# 'G'        soil heat flux (W m-2)
# 'EL'       altitude above sea level (m)
# 'PET'      daily potential evapotranspiration (mm)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     day of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

t_a <- dat$Ta       # daily mean air temperature
Rn  <- dat$Rn * 24 * 3600/1e6
                    # net solar radiation (convert W m-2 to MJ m-2 d-1)
G   <- dat$G * 24 * 3600/1e6
                    # soil heat flux  (convert W m-2 to MJ m-2 d-1)
EL  <- 288          # altitude

# Step 3: run the model
PET <- m6.PET.PriestleyTaylor(t_a, EL, G, Rn, alpha = 1.26)

# Step 4: present and visualize results
print(head(data.frame(
  t_a, Rn, G, EL, PET
)), digits = 3)     # summarize the result

plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Priestley-Taylor Model"
)                   # plot PET versus time
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv` and specify longitude and latitude of the study site.

```{r comment=""}
dat <- read.csv("./data/4-1A Daily Micrometeorological Measurements of Kellogg Biological Station.csv")
              # load monthly micrometeorological data set
              #     Column 1 ($Day):     day of the year
              #     Column 2 ($ET):      evapotranspiration (ET, W m-2)
              #     Column 3 ($wnd_spd): wind speed (m s-1)
              #     Column 4 ($Rn):      net solar radiation (W m-2)
              #     Column 5 ($VPD):     vapor pressure deficient (kPa)
              #     Column 6 ($Ta):      daily mean air temperature (Celsius)
              #     Column 7 ($h):       relative humidity (0-1)
              #     Column 8 ($VWC):     variable water content (0-1)
              #     Column 9 ($G):       soil heat flux (W m-2)

head(dat)
```

Extract micrometeorological variables that are required in Priestley-Taylor model from variable `dat`.

```{r comment=""}
t_a <- dat$Ta       # daily mean air temperature
Rn  <- dat$Rn * 24 * 3600/1e6
                    # net solar radiation (convert W m-2 to MJ m-2 d-1)
G   <- dat$G * 24 * 3600/1e6
                    # soil heat flux  (convert W m-2 to MJ m-2 d-1)
EL  <- 288          # altitude
```

Note that the factor `24 * 3600/1e6` is required for unit conversion from W m^-2^ to MJ m^-2^ d^-1^.

**Step3:** Run the model. The Priestley-Taylor PET model (Priestley and Taylor, 1972) is expressed as:

$$
\lambda\text{PET}=\alpha\frac{\Delta}{\Delta+\gamma}(R_n-G),\quad(1)
$$

where:

-   $\text{PET}$ = daily PET (mm d^-1^)

-   $\lambda$ = latent heat of vaporization (MJ kg^-1^), $\lambda=2.501-0.002361·T$

-   $T$ = daily mean air temperature (^o^C)

-   $\alpha$ = calibration constant, 1.26 for wet or humid conditions

-   $\Delta$ = slope of the saturation vapor pressure-temperature curve (kPa ◦C^-1^)

$$
\Delta = 0.100·(0.00738·T+0.8072)^7-0.000116,\quad(2)
$$

-   $\gamma$ = psychrometric constant modified by the ratio of canopy resistance to atmospheric resistance (kPa ◦C^-1^), $\gamma=c_pP/(0.622·\lambda)$

-   $c_p$ = specific heat of moist air at constant pressure (0.001013 MJ kg^−1^ ◦C^−1^)

-   $P$ = atmospheric pressure (kPa), $P=101.3-0.01055·EL$

-   $EL$ = elevation (m)

-   $G$ = heat flux density to the ground (MJ m^−2^ d^−1^)

-   $R_n$ = net radiation (MJ m^−2^ d^−1^)

```{r comment=""}
PET <- m6.PET.PriestleyTaylor(t_a, EL, G, Rn, alpha = 1.26)

print(head(data.frame(
  t_a, Rn, G, EL, PET
)), digits = 3)     # summarize the result
```

**Step4:** Visualize results. Create a line plot of predicted PET against time through the function `plot()`:

```{r comment=""}
plot(
  dat$Day, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Priestley-Taylor Model"
)                   # plot PET versus time
```

## References

Priestley, C. H. B., and Taylor, R. J. (1972). On the assessment of surface heat flux and evaporation using large-scale parameters. Monthly Weather Review, 100(2), 81--92.

Jasrotia, P., Xu, J., Bhardwaj, A. K., Hamilton, S. K., and Robertson, G. P. (2011). CO2 fluxes of transitional bioenergy crops: Effect of land conversion during the first year of cultivation. Global Change Biology-Bioenergy, 3(5), 401--412.
