---
title: "Calculate Potential Evapotranspiration Using Hargreaves-Samani Model"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate potential evapotranspiration (PET) using Hargreaves-Samani model.

```{r eval=FALSE}
# Calculate PET Using Hargreaves-Samani Model
# ********************************
#
# Calculate PET based on Hargreaves-Samani Model  (Priestley and Taylor， 1972).
# Data is from field measurements of micrometeorological variables in 2016 in an
# agricultural site with an eddy-covariance tower of the Kellogg Biological
# Station, Michigan, USA (Zenone et al. 2011).
#
# ********************************
# List of variables and units for this example
#
# 't_a'      daily mean air temperature (Celsius)
# 't_max'    daily maximum air temperature (Celsius)
# 't_min'    daily minimum air temperature (Celsius)
# 'Ra'       extraterrestrial radiation (MJ m-2 d-1)
# 'PET'      daily potential evapotranspiration (mm)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat <- read.csv("./data/1-1B Switchgrass Time Series.csv")
              # load monthly micrometeorological data set
              #   * column 1($date): observation date. Required format "dd/mm/yyyy"
              #   * column 2($t_max): daily maximum temperature (Celsius)
              #   * column 3($t_min: daily minimum temperature (Celsius)
lat <- 42.476708   # latitude of the site

t_max <- dat$t_max             # daily maximum air temperature (Celsius)
t_min <- dat$t_min             # daily minimum air temperature (Celsius)
t_a   <- (t_max + t_min)/2     # daily mean air temperature (Celsius)

## modeling extraterrestrial radiation
day   <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[1]))
              # extract the day number
month <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[2]))
              # extract the month number
year  <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[1]))
              # extract the year number
doys  <- DOY(year, month, day) # calculate day of the year
Ra    <- ra(lat, doys)         # calculate extraterrestrial radiation

# Step 3: run the model
PET <- m8.PET.HargreavesSamani(t_a, t_max, t_min, Ra)

# Step 4: present and visualize results
print(head(data.frame(
  t_a, t_max, t_min, Ra, PET
)), digits = 3)     # summarize the result

plot(
  doys, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Hargreaves-Samani Model"
)                   # plot PET versus time
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv` and specify latitude of the study site.

```{r comment=""}
dat <- read.csv("./data/1-1B Switchgrass Time Series.csv")
              # load monthly micrometeorological data set
              #   * column 1($date): observation date. Required format "dd/mm/yyyy"
              #   * column 2($t_max): daily maximum temperature (Celsius)
              #   * column 3($t_min: daily minimum temperature (Celsius)
lat <- 42.476708   # latitude of the site

head(dat)
```

Extract micrometeorological variables that are required in Hargreaves-Samani model from variable `dat`.

```{r comment=""}
t_max <- dat$t_max             # daily maximum air temperature (Celsius)
t_min <- dat$t_min             # daily minimum air temperature (Celsius)
t_a   <- (t_max + t_min)/2     # daily mean air temperature (Celsius)
```

Another variable required in Hargreaves-Samani is extraterrestrial radiation ($R_n$, MJ m^-2^ d^-1^). Although not measured in the provided data set, it can be calculated from latitude and day of the year (DOY). For more details, see Section 1-3.

```{r comment=""}
day   <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[1]))
              # extract the day number
month <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[2]))
              # extract the month number
year  <- as.numeric(lapply(strsplit(dat$date,"/"), function(xx) xx[1]))
              # extract the year number
doys  <- DOY(year, month, day) # calculate day of the year
Ra    <- ra(lat, doys)         # calculate extraterrestrial radiation

print(head(data.frame(DOY = doys, Ra = Ra)), digits = 3)
```

**Step3:** Run the model. The Hargreaves-Samani PET model (Hargreaves and Samani, 1982) is expressed as:

$$
\lambda·\text{PET}=0.0023·R_a·TD^{0.5}·(T+17.8),\quad(1)
$$

where:

-   $\text{PET}$ = daily PET (mm d^-1^)

-   $\lambda$ = latent heat of vaporization (MJ kg^-1^), $\lambda=2.501-0.002361·T$

-   $T$ = daily mean air temperature (^o^C)

-   $R_a$ = extraterrestrial solar radiation (MJ m^−2^ d^−1^)

-   $\text{TD}$ = daily differecnce between the maximum and minimum air temperature (^o^C)

```{r comment=""}
PET <- m8.PET.HargreavesSamani(t_a, t_max, t_min, Ra)

# Step 4: present and visualize results
print(head(data.frame(
  t_a, t_max, t_min, Ra, PET
)), digits = 3)     # summarize the result
```

**Step4:** Visualize results. Create a line plot of predicted PET against time through the function `plot()`:

```{r comment=""}
plot(
  doys, PET, type = "l", 
  xlab = "Day of the year", 
  ylab = "Daily PET",
  main = "PET Estimated by Hargreaves-Samani Model"
)                   # plot PET versus time
```

## References

Hargreaves, G. H., and Samani, Z. A. (1982). Estimating potential evapotranspiration. Journal of the Irrigation and Drainage Division, 108(3), 225--230.

Zenone, T., Chen, J., Deal, M. W., Wilske, B., Jasrotia, P., Xu, J., Bhardwaj, A. K., Hamilton, S. K., and Robertson, G. P. (2011). CO2 fluxes of transitional bioenergy crops: Effect of land conversion during the first year of cultivation. Global Change Biology-Bioenergy, 3(5), 401--412.
