---
title: "Calculate Evapotranspiration"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate evapotranspiration (ET) using empirical actual ET models.

## Code Example(s)

```{r eval=FALSE}
# Calculate ET Using Empirical Actual ET Models
# ********************************
#
# Calculate ET based on various empirical models. Data is from field measurements
# of micrometeorological variables in 2016 in an agricultural site with an 
# eddy-covariance tower of the Kellogg Biological Station, Michigan, USA (Zenone et
# al. 2011).
#
# ********************************
# List of variables and units for this example
#
# 'dat'          data set of micrometeorological variables:
#      * Column 1($month):      month of the year
#      * Column 2($LAI):        leaf area index
#      * Column 3($FAO.ETo):    monthly potential evapotranspiration (PET) estimted #                                    by FAO model (mm mo-1)
#      * Column 4($PET_Hamon):  PET estimated by Hamon's model (mm mo-1)
#      * Column 5($Priestley.TaylorPET):
#                               PET estimated by Priestley-Taylor model (mm mo-1)
#      * Column 6($P):          precipitation (mm mo-1)
#      * Column 7($VPD):        vapor pressure deficient (kPa)
#      * Column 8($Rn):         net solar radiation (W m-2)
#      * Column 9($SWC):        soil water content (%)
#      * Column 10($Ta):        mean air temperature (Celsius)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: read data
dat <- read.csv("./data/4-2 Calculating ET.csv")
dat$Rn <- dat$Rn * 30 * 24 * 3600 / 1e6 # unit conversion from W·m-2 to MJ·mo-1·m-2

# Step 3: run the model
## Method 1
ET_m1 <- with(dat, m1.ET.Sun(FAO.ETo, LAI, P))
plot(1:12, ET_m1, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 1)")

## Method 2
ET_m2 <- with(dat, m2.ET.Sun(HamonPET, LAI, P))
plot(1:12, ET_m2, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 2)")

## Method 3
ET_m3 <- with(dat, m3.ET.Fang(HamonPET, VPD, Rn))
plot(1:12, ET_m3, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 3)")

## Method 4
print(data.frame(
  Month = 1:12,
  ET_shurb = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P,land.type = "shrub")),
  ET_crop = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "crop")),
  ET_grass = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, P = P,land.type = "grass")),
  ET_deciduous = with(dat, m4.ET.type1(Rn = Rn, PET = HamonPET, land.type = "deciduous")),
  ET_needle = with(dat, m4.ET.type1(Rn = Rn, Ta = Ta, land.type = "needle")),
  ET_broad = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, Ta = Ta, land.type = "broad")),
  ET_mixed = with(dat, m4.ET.type1(PET= HamonPET, land.type = "mixed")),
  ET_savannas = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "savannas"))
), digits = 3)

## Method 5
print(data.frame(
  Month = 1:12,
  ET_shurb = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, P = P, land.type = "shrub")),
  ET_crop = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, P = P, land.type = "crop")),
  ET_grass = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, P = P, land.type = "grass")),
  ET_deciduous = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, land.type = "deciduous")),
  ET_broad = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, P = P, land.type = "broad")),
  ET_mixed = with(dat, m5.ET.type2(PET = HamonPET, land.type = "mixed")),
  ET_savannas = with(dat, m5.ET.type2(PET = HamonPET, LAI = LAI, P = P, land.type = "savannas"))
), digits = 3)

## Method 6
ET_m6 <- with(dat, m6.ET.Zhang(Priestley.TaylorPET, P, w=2))
plot(1:12, ET_m6, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 6)")
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Read data. Import the sample file through the function `read.csv`.

```{r comment=""}
dat <- read.csv("./data/4-2 Calculating ET.csv")
#      * Column 1($month):      month of the year
#      * Column 2($LAI):        leaf area index
#      * Column 3($FAO.ETo):    monthly potential evapotranspiration (PET) estimted #                                    by FAO model (mm mo-1)
#      * Column 4($PET_Hamon):  PET estimated by Hamon's model (mm mo-1)
#      * Column 5($Priestley.TaylorPET):
#                               PET estimated by Priestley-Taylor model (mm mo-1)
#      * Column 6($P):          precipitation (mm mo-1)
#      * Column 7($VPD):        vapor pressure deficient (kPa)
#      * Column 8($Rn):         net solar radiation (W m-2)
#      * Column 9($SWC):        soil water content (%)
#      * Column 10($Ta):        mean air temperature (Celsius)

print(dat)
```

Note that the units of net solar radiation is W m^-2^. Convert the unit into MJ mo^-1^ m^-2^.

```{r comment=""}
dat$Rn <- dat$Rn * 30 * 24 * 3600 / 1e6
```

**Step3:** Run the model. The ET model proposed by Sun et al. (2011a) is expressed as:

$$
ET = 11.94 + 4.76 · LAI + ET_o · (0.032 · LAI + 0.0026 · P + 0.15),\quad(1)
$$

where $ET_o$ is PET estimated by FAO method (See Section 4-1A), $LAI$ is leaf area index, $P$ is monthly precipitation (mm mo^-1^).

```{r comment=""}
ET_m1 <- with(dat, m1.ET.Sun(FAO.ETo, LAI, P))
plot(1:12, ET_m1, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 1)")
```

Another form of the ET model uses Hamon's PET model (Sun et al. 2011b):

$$ET = 0.174 · P + 0.502 · PET + 5.31 · LAI + 0.0222 · PET · LAI,\quad(2)$$

```{r comment=""}
ET_m2 <- with(dat, m2.ET.Sun(HamonPET, LAI, P))
plot(1:12, ET_m2, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 2)")
```

A further model proposed by Fang et al. (2016) is expressed as:

$$
ET = 0.42 + 0.74 · PET − 2.73 · VPD + 0.10 · R_n,\quad(3)
$$

where $PET$ is monthly potential ET (mm) estimated by Hamon's method. VPD (kPa) can be estimated from air temperature and relative humidity (See Section 1-2A, B). $R_n$ is net solar radiation (MJ mo^-1^ m^-2^).

```{r comment=""}
ET_m3 <- with(dat, m3.ET.Fang(HamonPET, VPD, Rn))
plot(1:12, ET_m3, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 3)")
```

Since $R_n$ is rarely available at the regional scale, another model that use precipitation $P$ is expressed as (Fang et al., 2016):

$$
ET = −4.79 + 0.75 · PET + 3.92 · LAI + 0.04 · P,\quad(4)
$$

Fang et al. (2016) further developed a collection of empirical ET models, referred as Type Ⅰ and Type Ⅱ, for different types of land covers. Type Ⅰ models are shown in the following table:

| Land cover type              | Model                                                                            |
|------------------------------|----------------------------------------------------------------------------------|
| Shrubland                    | $$                                                                               
                                                                ET = −4.59 + 13.02 · LAI + 0.10·R_n + 0.11 · P    
                                                                $$                                                |
| Cropland                     | $$                                                                               
                                                                ET = 0.87 + 0.19 · R_n + 13.99·LAI + 0.06 · P     
                                                                $$                                                |
| Grassland                    | $$                                                                               
                                                                ET = −0.87 + 0.20 · R_n + 0.10·P + 0.24 · SWC     
                                                                $$                                                |
| Deciduous forest             | $$                                                                               
                                                                ET = −14.22+0.74·PET+0.1·R_n                      
                                                                $$                                                |
| Evergreen needle leaf forest | $$                                                                               
                                                                ET = 13.47 + 0.10 · R_n + 1.35 · T_a              
                                                                $$                                                |
| Evergreen broad leaf forest  | $$                                                                               
                                                                ET = 0.01+0.63·T_a+0.46·SWC+0.14 · R_n            
                                                                $$                                                |
| Mixed forest                 | $$                                                                               
                                                                ET = −8.76 + 0.95 · PET                           
                                                                $$                                                |
| Savannas                     | $$                                                                               
                                                                ET = −8.07 + 33.46 · LAI + 0.07·R_n               
                                                                $$                                                |

where $T_a$ is the monthly mean air temperature (^o^C), $SWC$ is soil water content (%).

```{r comment=""}
print(data.frame(
  Month = 1:12,
  ET_shurb = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P,land.type = "shrub")),
  ET_crop = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "crop")),
  ET_grass = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, P = P,land.type = "grass")),
  ET_deciduous = with(dat, m4.ET.type1(Rn = Rn, PET = HamonPET, land.type = "deciduous")),
  ET_needle = with(dat, m4.ET.type1(Rn = Rn, Ta = Ta, land.type = "needle")),
  ET_broad = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, Ta = Ta, land.type = "broad")),
  ET_mixed = with(dat, m4.ET.type1(PET= HamonPET, land.type = "mixed")),
  ET_savannas = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "savannas"))
), digits = 3)
```

Type Ⅱ models are shown in the following table:

| Land cover type              | Model                                                                           |
|------------------------------|---------------------------------------------------------------------------------|
| Shrubland                    | $$                                                                              
                                                                ET = −3.11 + 0.39 ·PET+ 0.09·P + 11.127 · LAI    
                                                                $$                                               |
| Cropland                     | $$                                                                              
                                                                ET = −8.15 + 0.86 ·PET+ 0.01·P + 9.54 · LAI      
                                                                $$                                               |
| Grassland                    | $$                                                                              
                                                                ET =−1.36 + 0.70 · PET + 0.04·P + 6.56 · LAI     
                                                                $$                                               |
| Deciduous forest             | $$                                                                              
                                                                ET = −14.82+0.98·PET+2.72·LAI                    
                                                                $$                                               |
| Evergreen needle leaf forest | $$                                                                              
                                                                ET = 0.10 + 0.64 · PET + 0.04·P + 3.53 · LAI     
                                                                $$                                               |
| Evergreen broad leaf forest  | $$                                                                              
                                                                ET = 7.71 + 0.74 · PET + 1.85·LAI                
                                                                $$                                               |
| Mixed forest                 | $$                                                                              
                                                                ET = −8.763 + 0.95 · PET                         
                                                                $$                                               |
| Savannas                     | $$                                                                              
                                                                ET = −5.66 + 0.18 ·PET+ 0.10·P + 44.63 · LAI     
                                                                $$                                               |

```{r comment=""}
print(data.frame(
  Month = 1:12,
  ET_shurb = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P,land.type = "shrub")),
  ET_crop = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "crop")),
  ET_grass = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, P = P,land.type = "grass")),
  ET_deciduous = with(dat, m4.ET.type1(Rn = Rn, PET = HamonPET, land.type = "deciduous")),
  ET_needle = with(dat, m4.ET.type1(Rn = Rn, Ta = Ta, land.type = "needle")),
  ET_broad = with(dat, m4.ET.type1(Rn = Rn, SWC = SWC, Ta = Ta, land.type = "broad")),
  ET_mixed = with(dat, m4.ET.type1(PET= HamonPET, land.type = "mixed")),
  ET_savannas = with(dat, m4.ET.type1(Rn = Rn, LAI = LAI, P = P, land.type = "savannas"))
), digits = 3)
```

Zhang et al. (2001) developed a simple two-parameter ET model that can be used for examining the long-term average effect of vegetation changes on catchment ET:

$$
ET=P·\frac{1+w·\frac{PET}P}{1+w·\frac{PET}P+\frac{P}{PET}},\quad(5)
$$

where $w$ is the plant-available water coefficient that represents the relative difference in plant water use for transpiration. $PET$ is potential ET estimated by Priestley-Taylor model (See Section 4-1F). The best fitted w values for forest and grassland are 2.0 and 0.5, respectively.

```{r comment=""}
ET_m6 <- with(dat, m6.ET.Zhang(Priestley.TaylorPET, P, w=2))
  # argument 'w' is used to specify plant-available water coefficient
plot(1:12, ET_m6, type = "l", xlab = "Month", ylab = "ET",
     main = "ET estimates (Method 6)")
```

## References

Fang, Y., Sun, G., Caldwell, P., McNulty, S. G., Noormets, A., Domec, J. C., King, J., Zhang, Z., Zhang, X., Lin, G., Zhou, G., Xiao, J., Chen, J., and Zhou, G. (2016). Monthly land cover-specific evapotranspiration models derived from global eddy flux measurements and remote sensing data. Ecohydrology, 9(2), 248--266.

Sun, G., Alstad, K., Chen, J., Chen, S., Ford, C. R., Lin, G., Liu, C., Lu, N., McNulty, S. G., Miao, H., Noormets, A., Vose, J. M., Wilske, B., Zeppel, M., Zhang, Y., and Zhang, Z. (2011a). A general predictive model for estimating monthly ecosystem evapotranspiration. Ecohydrology, 4(2), 245--255.

Sun, G., Caldwell, P., Noormets, A., McNulty, S. G., Cohen, E., Moore Myers, J. M., Domec, J-C., Treasure, E., Mu, Q., Xiao, J., John, R., and Chen, J. (2011b). Upscaling key ecosystem functions across the conterminous United States by a watercentric ecosystem model. Journal of Geophysical Research: Biogeosciences, 116(G3), <https://doi.org/10.1029/2010JG001573.>

Zhang, L., Dawes, W. R., and Walker, G. R. (2001). Response of mean annual evapotranspiration to vegetation changes at catchment scale. Water Resources Research, 37(3), 701--708
