---
title: "Calculate GWP of Greenhouse Gases"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate global warming potential (GWP) of greenhouse gases (GHGs).

## Code Example(s)

```{r eval=FALSE}
# Calculate GWP of GHGs
# ********************************
#
# Calculate absolute (AGWP, W m-2 kg-1 yr) and relative GWP of three GHGs, namely
# carbon dioxide (CO2), methane (CH4), and nitrous oxide (N2O) for a period of time
#
# ********************************
require(bproce)

AGWP.CO2()                # AGWP of CO2 for 100 years
AGWP.CO2(TH = 20)         # AGWP of CO2 for 20 years

AGWP.CH4()                # AGWP of CH4 for 100 years
AGWP.CH4(TH = 20)         # AGWP of CH4 for 20 years
AGWP.CH4()/AGWP.CO2()     # GWP of CH4 for 100 years

AGWP.N2O()                # AGWP N2O for 100 years
AGWP.N2O(TH = 20)         # AGWP N2O for 20 years
AGWP.N2O()/AGWP.CO2()     # GWP N2O for 100 years
```

## Explanation

First we load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

The AGWP (W m^-2^ kg^-1^ yr) of a GHG pulse is calculates by integrating the radiative forcing (RF) due to GHG pulses over a time horizon:

$$
\text{AGWP}_x(\text{TH})=\int_0^\text{TH}\text{IRF}_x(t)·\text{mRF}_x/S,\quad(1)
$$

where $\text{TH}$ is the chosen time horizon (e.g. 20, 100, 500 years), $\text{mRF}$ is the marginal radiative forcing of a GHG species (a constant, W kg^-1^), $S$ is the surface area of the earth (m^2^), $\text{IRF}_x(t)$ is the impulse response function. For carbon dioxide (CO~2~), it is expressed as (IPCC, 2013, Skytt et al., 2020):

$$
\text{IRF}_{\text{CO}_2}(t)=0.2173+0.2240·\exp(\frac{-t}{394.4})+0.2824·\exp(\frac{-t}{36.54})+0.2763·\exp(\frac{-t}{4.304}),\quad(2)
$$

```{r comment=""}
cat(paste0(
  "AGWP of CO2 for 100 years = ", AGWP.CO2(), "\n",
  "AGWP of CO2 for 20 years = ", AGWP.CO2(TH = 20), "\n"
))
```

For methane (CH~4~) and nitrous oxide (N~2~O) , it is expressed as (IPCC, 2013):

$$
\text{IRF}_x(t)=\exp(-\frac{t}{\tau_x}),\quad(3)
$$

```{r comment=""}
cat(paste0(
  "AGWP of CH4 for 100 years = ", AGWP.CH4() , "\n",
  "AGWP of CH4 for 20 years = ", AGWP.CH4(TH = 20), "\n\n",
  "AGWP of N2O for 100 years = ", AGWP.N2O() , "\n",
  "AGWP of N2O for 20 years = ", AGWP.N2O(TH = 20), "\n"
))
```

The relative GWP, or simply GWP, is defined as the AGWP of a GHG species divided by the AGWP of CO~2~.

$$
\text{GWP}_x(\text{TH})=\frac{\text{AGWP}_x(\text{TH})}{\text{AGWP}_{\text{CO}_2}(\text{TH})},\quad(3)
$$

```{r comment=""}
cat(paste0(
  "GWP of CH3 for 100 years = ", round(AGWP.CH4()/AGWP.CO2(), 2), "\n",
  "GWP of N2O for 100 years = ", round(AGWP.N2O()/AGWP.CO2(), 2), "\n"
))
```

## References

IPCC (2013). Climate change 2013: The physical science basis. Contribution of Working Group I to the Fifth Assessment Report of the Intergovernmental Panel on Climate Change [Stocker, T.F., Qin, D. Plattner, G.-K., Tignor, M., Allen, S.K., Boschung, J., Nauels, A., Xia, Y., Bex V., & Midgley P. M. (eds.)]. Cambridge University Press, Cambridge, United Kingdom and New York, NY, USA, 1535 pp.

Skytt, T., Nielsen, S. N., and Jonsson, B. G. (2020). Global warming potential and absolute global temperature change potential from carbon dioxide and methane fluxes as indicators of regional sustainability---a case study of Jämtland, Sweden. Ecological Indicators, 110, 105831.

