---
title: "Calculating Upwelling Transmittance"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate upwelling transmittance value.

## Code Example(s)

```{r eval=FALSE}
# Calculate Upwelling Transmittance
# ********************************
#
# Calculate upwelling transmittance value from total incident sunslight and
# extraterrestrial radiation energy falling on a canopy surface.
#
# ********************************
# List of variables and units for this example
#
# 'year'        experimental year
# ’month'       experimental month
# 'day'         experimental day
# 'hour'        hour of the day
# 'lng'         longitute of the study site (decimal)
# 'lat'         latitude of the study site (decimal)
# 'Sw_down'     total incident sunlight (W m-2)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: choose parameters
year    <- 2018
month   <- 7
day     <- 27
hour    <- 9.4
lng     <- -85.446758
lat     <- 42.476708
Sw_down <- 691.55

## Step3: run the model
doy    <- DOY(year, month, day)
zenith <- zangle(year, month, day, lng, lat, hour)[[2]]
transmittance(Sw_down, doy, zenith)
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `day`, `month`, and `year`: Experimental day. The provided example calculates the upwelling transmittance value for 27 July 2018 and, therefore, we assigns 27to the variable `day`, 7 to `month`, 2018 to `year`.

-   `hour`: Hour of the day used to calculate the zenith angle. The provided example uses 9.4h (9: 24 a.m).

-   `lng`: Longitude of the study site (in decimal). The provided example uses 85.446758^o^W.

-   `lat`: Latitude of the study site (in decimal). The provided example uses 42.476708^o^N.

-   `Sw_down`: Total incident sunlight (W m^-2^). The provided example uses 691.55 W m^-2^.

```{r comment=""}
year    <- 2018
month   <- 7
day     <- 27
hour    <- 9.4
lng     <- -85.446758
lat     <- 42.476708
Sw_down <- 691.55
```

**Step3:** Run the model. The first step is to calculate day of the year (DOY) using the following equation:

$$
\text{DOY}=\left[ \frac{275\text{month}}{9} \right]-\left[\frac{\text{month}+9}{12}\right]·
\left(\left[ \frac{\text{year}-4\text[\text{year}/4]+2}{3} \right]+1\right)+\text{day}-30,\quad(1)
$$

```{r comment=""}
doy <- DOY(year, month, day)
cat(paste0(day,"/",month,"/",year," is the ",doy,"th day of ",year,".\n"))
```

Then calculate the zenith angle (For details, see Section 1-5A, B):

```{r comment=""}
zenith <- zangle(year, month, day, lng, lat, hour)[[2]]
cat(paste0(
  "Date: ", day,"/",month,"/",year, "\n",
  "Hour: ", hour, "\n",
  "Zenith Angle: ", round(zenith, 3), "\n"
))
```

The transmittance value is calculated as:

$$
T_a=\frac{Sw_↓}{\text{SW}_{\text{TOA}}},\quad(2)
$$
where $Sw_↓$ is total incident sunlight (W m^-2^), $\text{SW}_{\text{TOA}}$ is the value of extraterrestrial radiation energy falling on a canopy surface (W m^-2^), which is calculated as:

$$
\text{SW}_{\text{TOA}}=I_{sc}·d_r·I_{\theta},\quad(3)
$$
where $I_{sc}$ is the solar constant (1367 W m^-2^), $I_\theta$ is the extraterrestrial radiance intensity onto a crop canopy:

$$
I_{\theta}=\cos(\theta_{\text{ZENITH}}·\frac{\pi}{180}),\quad(4)
$$

where $\theta_{\text{ZENITH}}$ is zenith angle. $d_r$ is the earth-sum distance:

$$
d_{r}=1+0.0334\cos(\text{DOY}·\frac{2\pi}{365.25}),\quad(5)
$$

```{r comment=""}
Ta <- transmittance(Sw_down, doy, zenith)
cat(paste0(
  "Date: ", day,"/",month,"/",year, "\n",
  "Hour: ", hour, "\n",
  "Incident Sunlight: ", Sw_down, "\n",
  "Transmittance : ", round(Ta, 3), "\n"
))
```
