---
title: "Calculate GWP from Surface Albedo"
author: "Hao-Ran Wu ( haoranwu @ zju.edu.cn )"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section we are concerned about how to calculate global warming potential (GWP) from surface albedo.

## Code Example(s)

```{r eval=FALSE}
# Calculate GWP from Surface Albedo
# ********************************
#
# Calculate GWP from surface albedo change.
#
# ********************************
# List of variables and units for this example
#
# 'Sw_up'         reflected light (W m-2)
# 'Sw_down'       total incident sunlight (W m-2)
# 'alpha_REF'    	surface albedo of the reference site (unitless)
# 'Ta'            upwelling transmittance value (unitless)
#
# ********************************
# Step 1: import package
require(bproce)

# Step 2: Choose parameters
Sw_up     <- 134
Sw_down   <- 478
alpha_REF <- 0.172
Ta        <- 0.854

# Step 3: Run the model
GWP.albedo(Sw_up, Sw_down, alpha_REF, Ta)
```

## Explanation

**Step1:** Load the package `bproce` through the `require()` function.

```{r message=FALSE}
require(bproce)
```

**Step2:** Choose model parameters.

-   `Sw_up`: Reflected light (W m^-2^). The provided example uses 134 W m^-2^.

-   `Sw_down`: Total incident sunlight (W m^-2^). The provided example uses 478 W m^-2^.

-   `alpha_REF`: Surface albedo of the reference site. The provided example uses 0.172.

-   `Ta`: Upwelling transmittance value (See Section 5-2 for how to calculate). The provided example uses 0.854.

```{r comment=""}
Sw_up     <- 134
Sw_down   <- 478
alpha_REF <- 0.172
Ta        <- 0.854
```

**Step3:** Run the model. Albedo is defined as the ratio of reflected light ($Sw_↑$, W m^-2^) to the total incident sunlight ($Sw_↓$↑$, W m^-2^) for a given area of land surface. GWP from surface albedo can be calculated as:

$$
\text{GWP}=\frac{\Delta\text{RF}_{\text{TOA}}}{\text{AF}_{\text{CO}_2}·\text{mRF}_{\text{CO}_2}}·\frac 1 {\text{TH}},\quad(1)
$$

where $\text{AF}_{\text{CO}_2}$ is the percentage of emitted CO~2~ that can be calculated from impulse response function (IRF, See Section 5-1), $\text{TH}$ is the time horizon for 100 years ($\text{TH} = 100$), $\text{mRF}_{\text{CO}_2}$ is marginal radiative forcing of CO~2~ (a constant, W m^-1^), $\Delta{RF}_{\text{TOA}}$ is the radiative forcing from changes in albedo (W m^-1^), which can be calculated via IRD (equation (2)) or AGWP (equation (3)) method:

$$
\Delta\text{RF}_{\text{TOA}}=-\frac1NT_a·\text{RF}_{\Delta\alpha},\quad(2)
$$

$$
\Delta\text{RF}_{\text{TOA}}=-\frac1N\frac{T_a·\text{RF}_{\Delta\alpha}}{S_{\text{Earth}}},\quad(3)
$$

where $N$ is the number of days (or hours), $T_a$ is the upwelling transmittance value (See Section 5-2 for calculation), $S_{\text{Earth}}$ is the surface area of the earth, $\text{RF}_{\Delta\alpha}$ is the change in net radiative forcing from the surface driven by surface albedo (W m^-1^), which is can be calculated as:

$$
\text{EF}_{\Delta\alpha}=-Sw_↓\Delta\alpha_s,\quad(4)
$$

where $Sw_↓$ is local incoming solar radiation incident to the surface (W m^-2^), $\Delta\alpha_s$ is change in albedo , defined by the difference between the albedo value of the target ecosystem and the reference site.

```{r comment=""}
gwp <- GWP.albedo(Sw_up, Sw_down, alpha_REF, Ta)
cat("GWP (relative to carbon dioxide) from surface albedo =", round(gwp, 3), "\n")
```
